---
title: "mini-project-2"
author: "Ziqi Su"
date: "10/10/2020"
output: md_document
---

**Version 1.1**

*To complete this milestone, edit this `.rmd` file directly. Fill in the sections that are commented out with `<!--- start your work here--->`. When you are done, make sure to knit to an `.md` file by changing the YAML settings, before submitting a tagged release on canvas.*

# Welcome back to your mini data analysis project!

This time, we will explore more in depth the concept of *tidy data*, and hopefully investigate further into your research questions that you defined in milestone 1.

Begin by loading your data and the tidyverse package below: 

```{r, message = FALSE}
 #install.packages("tsibble")
 #install.packages("lubridate")
library(datateachr) # <- contains the data you picked!
library(tidyverse)
library(lubridate)
```

# Learning Objectives

By the end of this milestone, you should:

+ Become familiar with manipulating and summarizing your data in tibbles using `dplyr`, with a research question in mind
+ Understand what *tidy* data is, and how to create it. In milestone 3, we will explore when this might be useful
+ Generate a reproducible and clear report using R Markdown
+ Gain a greater understanding of how to use R to answer research questions about your data

**Things to keep in mind**

+ Remember to document your code, be explicit about what you are doing, and write notes in this markdown document when you feel that context is required. Create your analysis as if someone else will be reading it! **There will be 2.5 points reserved for reproducibility, readability, and repo organization.**

+ Before working on each task, you should always keep in mind the specific **research question** that you're trying to answer.

# Task 1: Process and summarize your data (15 points)

From milestone 1, you should have an idea of the basic structure of your dataset (e.g. number of rows and columns, class types, etc.). Here, we will start investigating your data more in-depth using various `dplyr` functions. 

### 1.1 (2.5 points) 

First, write out the 4 research questions you defined in milestone 1 were. This will guide your work through milestone 2:

<!-------------------------- Start your work below ---------------------------->
1. Question1: What's the trend of the number of published games in recent years?

2. Question2: Which genre does customers prefer? (i.e. the distribution of the number of positive reviews vs genres)

3. Question3: If a game is discounted, is it more likely to be a "bad quality" game, in another word, do discounted games received more negative reviews comparing original-price games?

4. Question4: How many achievements in a game do customers prefer?

<!----------------------------------------------------------------------------->

### 1.2 (10 points)

Now, for each of your four research questions, choose one task from options 1-4 (summarizing), and one other task from 4-8 (graphing). You should have 2 tasks done for each research question (8 total). Make sure it makes sense to do them! (e.g. don't use a numerical variables for a task that needs a categorical variable.). Comment on why each task helps (or doesn't!) answer the corresponding research question. 

Ensure that the output of each operation is printed! 

**Summarizing:**

1. Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.
2. Compute the number of observations for at least one of your categorical variables. Do not use the function `table()`!
3. Create a categorical variable with 3 or more groups from an existing numerical variable. You can use this new variable in the other tasks! *An example: age in years into "child, teen, adult, senior".*
4. Based on two categorical variables, calculate two summary statistics of your choosing.

**Graphing:**

5. Create a graph out of summarized variables that has at least two geom layers.
6. Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are "pretty" or easier to read. 
7. Make a graph where it makes sense to customize the alpha transparency.
8. Create 3 histograms out of summarized variables, with each histogram having different sized bins. Pick the "best" one and explain why it is the best.

Make sure it's clear what research question you are doing each operation for!

<!------------------------- Start your work below ----------------------------->
+ q1. What's the trend of the number of published games in recent years?

First I need to compute how many games were published in each year, corresponding to task 2 for *summarizing*. It is a bit tricky because the dates are given as strings, so I have to extract the years and  store them into int before starting works.
```{r}
nperyear <-steam_games %>%
  filter(!is.na(release_date)& as.matrix(release_date)!="NaN")%>%  #remove all rows with NA or NaN values
   mutate(year=strtoi(unlist(strsplit(release_date,","))[2*(1:length(release_date)) ])) %>%#Since all date are given by string, in form of "May 12, 2016". I first split the date into 2 strings separated by ",", then store the even to a new variable "year".
   filter(year<=2019)%>%  #Since games published after 2020 maybe not well collected, so I filter out all games after 2020
   group_by(year)%>%
   summarise(n=n())
nperyear
```
Second, I'd like to plot a graph to show the relationship between years and the number of published games each year. By graphing out the trend, I found that the increment over years is exponential, so it makes sense to scale the number of published games by *log10*, corresponging to task 6 in *graphing*.
```{r}
 nperyear %>%
   ggplot(aes(x=year, y=n)) +
   geom_point() +
   scale_y_log10() +
   ylab("The Number of Published Games (log10)")
```
 +q2.  Which genre does customers prefer? (i.e. the distribution of the number of positive reviews vs genres)
 
 ++ First I need to create a categorical variable named *tone* to store the tone of reviews from variable *all_reviews*. There will be three groups: "very-positive(more than 80% positive reviews)", "mostly positive (more than 80% positive reviews)" and "Mixed reviews (more than 80% positive reviews)". It's corresponding to task 3 in *summarizing*, although the *all_reviews* variable is not a numerical variable, the same process need to be done here so I believe the work here is close to the task 3.
```{r}
tones <-steam_games %>%
# remove all lines that is "NA" or "NaN" in all_reviews
  filter(!is.na(all_reviews)& as.matrix(all_reviews)!="NaN") %>% 
   select(name,genre,all_reviews)%>% 
# There are 3 groups based on the *all_reviews* variable
   mutate(
    tone = case_when(
      str_detect(all_reviews, "Very Positive") ~ "Very Positive",
      str_detect(all_reviews, "Mostly Positive") ~ "Mostly Positive",
      str_detect(all_reviews, "Mixed") ~ "Mixed"
    )
  )
tones
```

++ Then, I need to plot the relationship between tones and genres. I set the tones as the factor and genre as variable, I could be able to plot it in the way that use 3 geom_bar layers to plot three tones one by one, but this code does the same thing, so I assume this is close to task 5 in *graphing*.

```{r}
 genre_reviews <-tones %>%
   filter(!str_detect(genre,",")& ! is.na(tone)) %>% # remove games that have more thna one genres
   group_by(genre,tone) 
genre_reviews%>%
   ggplot(aes(factor(tone))) +
  geom_bar(aes(x=genre, fill = tone, group = tone), 
           position = "stack") +
  scale_fill_discrete("tone") +
  labs(y = "Proportion", x = "Cylinders") +
  theme_minimal() +
   theme(axis.text.x = element_text(angle=90))  # Rotate texts on X-axis to avoiding texts overlapping.
```

+ q3 : If a game is discounted, is it more likely to be a "bad quality" game, in another word, do discounted games received more negative reviews comparing original-price games? 

++ To answer this question, first, I need to compute the *range*, *mean*, *median* and *quantile (0.1,0.9)* across years as task 4 of *summarizing*.

```{r}
games_withyear <-steam_games %>%    #Explained before
  filter(!is.na(release_date)& as.matrix(release_date)!="NaN")%>%  #remove all rows with NA or NaN values
   mutate(year=strtoi(unlist(strsplit(release_date,","))[2*(1:length(release_date)) ])) %>%#Since all date are given by string, in form of "May 12, 2016". I first split the date into 2 strings separated by ",", then store the even to a new variable "year".
   filter(year<=2019)

discounted<-games_withyear %>% 
    filter( !is.na(original_price)  & as.matrix(original_price)!=0)  %>% #remove games those missing price information
    filter(as.matrix(discount_price)<=as.matrix(original_price))  %>%
   group_by(year) %>%
   summarise(mean_original = mean(original_price),
             min_original=min(original_price), 
             max_original=max(original_price),
             median_original=median(original_price),
             qs_original = quantile(original_price, c(0.25, 0.75)),
             mean_discount = mean(discount_price))
discounted
```

 ++ To visualize the means of original-prices and discount-prices over years, I can plot them in a graph, which correspongding to task 5 in *graphing*.
```{r}
(discounted %>%
   ggplot()+
   geom_point(aes(x=year,y=mean_original))  +
   geom_jitter(aes(x=year,y=mean_discount),color='red'))

```

 + q4: How many achievements in a game do customers prefer?
 ++ First, I need to have a brief exploration of the "achievements" variables, I am going to compute the *range*, *mean*, and *median* and *quantile (0.1,0.9)* across different games(i.e. game *name* as categorical variable) as task 4 of *summarizing*.
```{r}
(achievement_stat <- steam_games %>%
   select(name, achievements)%>%
  filter(!is.na(achievements)& as.matrix(achievements)!="NaN")%>%  #remove all rows with NA or NaN element
   summarise(mean = mean(achievements),
             min=min(achievements), 
             max=max(achievements),
             median=median(achievements),
             qs = quantile(achievements, c(0.1, 0.9)) ))
```
With the output, I know that the average number of achievements is around 77, while most games have 12-38 achievements. Some rare cases (i.e. a game has more than thousands of achievements) raised up the average number of achievements.

 ++ To visualize the achievements distribution, I will graph histogram. As required in task 8, I will create 3 histograms out of summarized variables with binwidth =1, 10, 20.
```{r}
# binwidth=1
steam_games %>%
   filter(achievements<100)%>% #Since the 95% of the number of achievements are under 100,so I removed all rows that have more than 100 achievements.
   ggplot(aes(x=achievements)) +
   geom_histogram(binwidth = 1) 
```
```{r}
# binwidth=10
steam_games %>%
   filter(achievements<100)%>% #Since the 95% of the number of achievements are under 100,so I removed all rows that have more than 100 achievements.
   ggplot(aes(x=achievements)) +
   geom_histogram(binwidth = 10) 
```

```{r}
# binwidth=20
steam_games %>%
   filter(achievements<100)%>% #Since the 95% of the number of achievements are under 100,so I removed all rows that have more than 100 achievements.
   ggplot(aes(x=achievements)) +
   geom_histogram(binwidth = 20) 
```
We could see binwidth=1 is too small so that it shows more information than needed, and binwidth=20 is too large so that it ignores a lot details.


<!----------------------------------------------------------------------------->

### 1.3 (2.5 points)

Based on the operations that you've completed, how much closer are you to answering your research questions? Think about what aspects of your research questions remain unclear. Can your research questions be refined, now that you've investigated your data a bit more? Which research questions are yielding interesting results?

<!------------------------- Write your answer here ---------------------------->
+ How much closer are you to answering your research questions?

My results well answered my questions one, two and fur. The result of question three still cannot answer it, because the resulted graph can only show what portions of price that games discounted, instead of showing the relationships between reviews and discounted games. Further works need to be done to reveal those relationships.

+ What aspects of your research questions remain unclear?
 
 The relationship between reviews and original/discounted games is still unclear. The variable *discount price* itself is a bit ambiguous: *NA* in *discount price* may represent that the game didn't have a discount, also could represent a missing data; and even some discount_prices are higher than their original_prices, so the comparison between original-price and discount-price cannot clearly show discounting information. 
 
 +Now can your research questions be refined?
 
 I'd like to refine question three to avoid using the variable *discount price*. Maybe a question like "Is there any relationship between price and the number of achievement" will be much more clear to answer.
 
 +Any interesting findings?
 
 Some of my questions are asking about the relationships between reviews and a feature of games.There might be many causes resulting in a negative reviews, so the trends or relationships I found above might be misled by some undiscovered reasons.

<!----------------------------------------------------------------------------->

# Task 2: Tidy your data (12.5 points)

In this task, we will do several exercises to reshape our data. The goal here is to understand how to do this reshaping with the `tidyr` package.

A reminder of the definition of *tidy* data:

- Each row is an **observation**
- Each column is a **variable**
- Each cell is a **value**

*Tidy'ing* data is sometimes necessary because it can simplify computation. Other times it can be nice to organize data so that it can be easier to understand when read manually. 

### 2.1 (2.5 points)

Based on the definition above, can you identify if your data is tidy or untidy? Go through all your columns, or if you have >8 variables, just pick 8, and explain whether the data is untidy or tidy.

<!--------------------------- Start your work below --------------------------->
Based on my research questions, I will choose 8 related columns for tasks in this part.
```{r}
(eight_columns <-steam_games %>%
   select(id,name,genre, achievements,release_date,developer,original_price,discount_price))
```

My data is untidy:

   In *genre* column, each cell might contain more than one "value", for example, if a game belongs to *action*, *adventure* and *strategy*, its cell will store "Action,Adventure,Strategy".
<!----------------------------------------------------------------------------->

### 2.2 (5 points)

Now, if your data is tidy, untidy it! Then, tidy it back to it's original state.

If your data is untidy, then tidy it! Then, untidy it back to it's original state.

Be sure to explain your reasoning for this task. Show us the "before" and "after".

<!--------------------------- Start your work below --------------------------->
+ Before tidying:
```{r}
print(eight_columns,options(pillar.min_chars = Inf))
```
 I set"options(pillar.min_chars = Inf)" to print the entire string in each column, it works well when I ran it in console, however, when I output it as .md file, it cannot be show as entire string. I have no idea why the console and markdown file give different results, I searched online and somebody said it is related to Rstudio environment and profiles.
 So I manually show it below. I know it's not reproducible but I believe this problem is out of the scope of STAT545.
 
id        name                                           genre                   
<dbl>    <chr>                                           <chr>                                                                
1           DOOM                                         Action                                                     
2           PLAYERUNKNOWN'S BATTLEGROUNDS                Action,Adventure,Massively Multiplayer                          
3           BATTLETECH                                   Action,Adventure,Strategy         

We can see in column *genre* of rows 2,3,4 and 5,the cells contain more than a "value", which makes it not a tidy data.

+ After tidying:
```{r}
tidy_data1 <- eight_columns %>%
   separate_rows(genre, convert = TRUE) # separate_rows is a function included in tidyr
print(tidy_data1,options(pillar.min_chars = Inf))
```
id        name                                           genre                   
<dbl>    <chr>                                           <chr>                                                                
1           DOOM                                         Action                                             
2           PLAYERUNKNOWN'S BATTLEGROUNDS                Action
2           PLAYERUNKNOWN'S BATTLEGROUNDS                Adventure
2           PLAYERUNKNOWN'S BATTLEGROUNDS                Massively Multiplayer            
3           BATTLETECH                                   Action 
3           BATTLETECH                                   Adventure  
3           BATTLETECH                                   Strategy   

We can see that those cells are separated into multiple rows. Now, each cell contains only one "value", which make data tidier than before.

To untidy it back to original state:

```{r}
tidy_data2 <- tidy_data1 %>%
   group_by(id,name, achievements,release_date,developer,original_price,discount_price) %>% 
   summarise(genre  = paste(genre, collapse =","))
print(tidy_data2,options(pillar.min_chars = Inf))
```
The tibble here is untidied back to original state (the *genre* column is placed at the most right side).

<!----------------------------------------------------------------------------->

### 2.3 (5 points)

Now, you should be more familiar with your data, and also have made progress in answering your research questions. Based on your interest, and your analyses, pick 2 of the 4 research questions to continue your analysis in milestone 3, and explain your decision. 

Try to choose a version of your data that you think will be appropriate to answer these 2 questions in milestone 3. Use between 4 and 8 functions that we've covered so far (i.e. by filtering, cleaning, tidy'ing, dropping irrelvant columns, etc.). Save this tibble as an `.RDS` object with the function `saveRDS()`. 

<!--------------------------- Start your work below --------------------------->
I 'd like to choose question one and two to continue your analysis in milestone 3.
 + Question1: What's the trend of the number of published games in recent years?

 + Question2: Which genre does customers prefer? 

I make my_data as the version of dataset that I will use in milestone 3.
```{r}
my_data <-steam_games %>%
   filter(!is.na(all_reviews) &!is.na(release_date)& as.matrix(release_date)!="NaN")%>%  #remove all rows with NA or NaN values
   separate_rows(genre, convert = TRUE) %>%
   mutate(tone = case_when(str_detect(all_reviews, "Very Positive") ~ "Very Positive",
                           str_detect(all_reviews, "Mostly Positive") ~ "Mostly Positive",
                           str_detect(all_reviews, "Mixed") ~ "Mixed") ) %>%
   select(id,name,genre, tone,release_date,achievements,developer,original_price,discount_price) 
saveRDS(my_data,file="my_data")
```

<!----------------------------------------------------------------------------->

*When you are done, knit an `md` file. This is what we will mark! Make sure to open it and check that everything has knitted correctly before submitting your tagged release.*